

---
output:
  html_document: default
  pdf_document: default
---

##Title: "Statistical Inference Course Project Part 1"##
**author:**"Sohair Zaki"
**date:**"May 12, 2017"

Part 1: Simulation Exercise 
In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials. We need to do a thousand simulations.
pseodocode
Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

--Show the sample mean and compare it to the theoretical mean of the distribution.
--Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
--Show that the distribution is approximately normal. 
---
start r code
```{r}

# 1. Show the sample mean and compare it to the theoretical mean of the distribution.
    # Set seed
    set.seed(2)

    # Set lambda per value provided 
    lambda <- 0.2

    # Simulations 1000 simulation
    sim <- 1000

    # Samples we are given 40 exponentials
    n <- 40
    
    #exponential distribution can be simulated in r using rexp 
    # we need to replicate that 1000 times per requirements 1000 simulations
    
    ExpSimulation <-  replicate(sim, rexp(n,lambda))
    
    #take the mean 
    ExpMean <- colMeans (ExpSimulation)
    
    # simulated mean
    SimuMean <- mean (ExpMean)
    
    # let us display it 
    SimuMean
    
    #note it is equal to 5.016356
    
    # Therorical mean is 1/lambda
    TheroticalMean <- 1/0.2
    #let us display it
    TheroticalMean
    # note it is equal to 5
 # conclusion : The simulated mean is 5.01 and  theoretical mean is 5.   
 ```

```{r, fig.width=7, fig.height=6}
# plot size is 7 inches wide and 6 inches high
 # let us draw a histogram for the mean of the simulated exponential ==ExpMean
# xlab=label of x axis is Mean of Exponential 
# main= title of the graph Histogram of 1000 simualtions 
# col= color of the bars 
hist(ExpMean, breaks = 20, xlab = "Mean of Exponential", 
     main = "Histogram of 1000 Simulated Exponential", 
     col = "lightblue")
    
    
```
```{r}
#Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
# get the standard deviation of ExpMean using sd()

SimuSTD <- sd(ExpMean)
# get its variance 

SimuVAR <- SimuSTD^2

# let us display it
SimuVAR

#Note the result is 0.6691305

# Theoretical Standard Deviation is 1/lambda/sqrt (n)
TheroticalSTD <- (1/lambda)/sqrt(n)
# Theoretical variance
TheroticalVAR <- TheroticalSTD^2

# let us display it 
TheroticalVAR
# the result is 0.625

# conclusion The simulated variance is 0.6691 cand theoretical variance of 0.625.

#3. Show that the distribution is approximately normal
# Plot Histogram
```
```{r, fig.width=7, fig.height=6}
# plot size is 7 inches wide and 6 inches high
 # let us draw a histogram for the mean of the simulated exponential ==ExpMean
# xlab=label of x axis is Mean of Exponential 
# main= title of the graph Histogram of 1000 simualtions 
# col= color of the bars 
hist(ExpMean, breaks = 40, xlab = "Mean of Exponential", 
     main = "Histogram of 1000 Simulated Exponential", 
     col = "lightblue")

# Add the Theoretical Normal Distribution Line
# generate a seq 
x  <- seq(min(ExpMean), max(ExpMean), length = 100) 

y  <- dnorm(x , mean = 1/lambda, sd = 1/lambda/sqrt(n))
# draw lines with line type 6


lines(x , y *100, lty=6, lwd=2)

```
Conclusion :the distribution is closely matched with a normal distribution.

